An Analysis of Health & Lifestyle Factors Associated with Sleep Quality

### 2024-05-01 R/analysis-code/statistical-analysis.qmd
References Check

library(broom) 
library(here) 
library(tidymodels)
library(tidyverse)
library(corrplot)
library(ranger)
library(vip)
library(ggplot2)
library(flextable)
library(gt)
```

@@

####Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
```{r}
lmfit_quality_duration <- lm(Quality.of.Sleep ~ Sleep.Duration, sleepdata)

@@

#### Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors
```{r}
lmfit_quality_duration_occupation <- lm(Quality.of.Sleep ~ Sleep.Duration + Occupation, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Blood Pressure category as a predictor.
```{r}
lmfit_quality_BP <- lm(Quality.of.Sleep~ cat_bp, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.
```{r}
lmfit_quality_stress <- lm(Quality.of.Sleep~ Stress.Level, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Age and Gender as predictors.
```{r}
lmfit_quality_gender_age <- lm(Quality.of.Sleep~ Gender + Age, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Age as a predictor.
```{r}
lmfit_quality_age <- lm(Quality.of.Sleep~ Age, sleepdata)

@@

Now we'll create a new variable that groups the different occupations to make analysis more succinct.
```{r}
# Create a new column

@@

We'll make sure there isn't collinearity between our continuous variables before we fit more complex models.
```{r}
# Select the variables

@@

First, we'll create a random seed to aid in reproducibility.
```{r}
#Set the seed
rngseed=1234

@@

Now, we'll make a "null" model and compare its results to that of our RF in order to make sure that the RF is useful.
```{r}
# Specify the null model

@@

null_fit <- fit(null_mod, Quality.of.Sleep ~ 1, data = sleepdatafinal)

@@

Now we'll make a table for the manuscript comparing the metrics we evaluated.

```{r, message=FALSE}
#extract the metric values for the null model from the objects they're stored in

null_rmse_value <- null_rmse$.estimate

null_r_squared_value <- r_squared$.estimate

# Create a data frame

results <- data.frame(

Model = c("Random Forest", "Null Model"),

`R-squared` = c(rf_r_squared, null_r_squared_value),

RMSE = c(rf_rmse, null_rmse_value)

# Create the table

gt_table <- gt(results)

print(gt_table)
```

>>>>

library(broom) 
library(here) 
library(tidymodels)
library(tidyverse)
library(corrplot)
library(ranger)
library(vip)
library(ggplot2)
library(flextable)
library(gt)
library(glmnet)
```

@@

####Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor

```{r}
lmfit_quality_duration <- lm(Quality.of.Sleep ~ Sleep.Duration, sleepdata)

@@

#### Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors

```{r}
lmfit_quality_duration_occupation <- lm(Quality.of.Sleep ~ Sleep.Duration + Occupation, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Blood Pressure category as a predictor.

```{r}
lmfit_quality_BP <- lm(Quality.of.Sleep~ cat_bp, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.

```{r}
lmfit_quality_stress <- lm(Quality.of.Sleep~ Stress.Level, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Age and Gender as predictors.

```{r}
lmfit_quality_gender_age <- lm(Quality.of.Sleep~ Gender + Age, sleepdata)

@@

####Fitting a model using Sleep Quality as the outcome and Age as a predictor.

```{r}
lmfit_quality_age <- lm(Quality.of.Sleep~ Age, sleepdata)

@@

Now we'll create a new variable that groups the different occupations to make analysis more succinct.

```{r}
# Create a new column

@@

We'll make sure there isn't collinearity between our continuous variables before we fit more complex models.

```{r}
# Select the variables

@@

First, we'll create a random seed to aid in reproducibility.

```{r}
#Set the seed
rngseed=1234

@@

Now, we'll make a "null" model and compare its results to that of our RF in order to make sure that the RF is useful.

```{r}
# Specify the null model

@@

null_fit <- fit(null_mod, Quality.of.Sleep ~ ., data = sleepdatafinal)

@@

# Fitting a LASSO Model using cross validation

First, we'll create a random seed to aid in reproducibility.

```{r}
#Set the seed
rngseed= 3456

We'll now train a LASSO model using CV with no train/test split (as the data has less than 400 observations and many of the values are rare/unique for some columns). We'll follow the tidymodels framework.

```{r}
# Recipe preparation
data_recipe <- recipe(Quality.of.Sleep ~ ., data = sleepdatafinal) %>%
  step_dummy(all_nominal(), -all_outcomes())  # One-hot encode categorical variables

# Prepare the recipe
prepared_recipe <- prep(data_recipe)

# Apply the prepared recipe to transform your data
processed_data <- bake(prepared_recipe, new_data = sleepdatafinal)

# Define the LASSO model
lasso_model <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

# Set up a grid of values to tune over
lasso_grid <- tibble(penalty = 10^seq(-5, 2, length.out = 50))

# Create a workflow
lasso_wf <- workflow() %>%
  add_recipe(data_recipe) %>%
  add_model(lasso_model)

resamples2 <- vfold_cv(sleepdatafinal, v = 5, repeats = 5)

# Tune the LASSO model using cross-validation
lasso_tune <- tune_grid(
  lasso_wf,
  resamples = resamples2,  
  grid = lasso_grid,
  metrics = metric_set(rmse), 
  control = control_grid(save_pred = TRUE)
)

# Plot results
autoplot(lasso_tune)

# Select the best hyperparameters
best_param <- select_best(lasso_tune, metric = "rmse")

# Finalize the workflow with the best hyperparameters
final_wf2 <- finalize_workflow(lasso_wf, best_param)

# Fit the final LASSO model using the entire dataset
final_fit2 <- fit(final_wf2, data = sleepdatafinal)

# Extract the fitted LASSO model
fitted_model2 <- extract_fit_parsnip(final_fit2)

# Create the variable importance plot
vip(fitted_model2$fit)

```
*Kevin: Please check this*

#### Predicting on new data

predictions2 <- predict(final_fit2, new_data = sleepdatafinal)

sleepdatafinal$Predicted <- predictions2$.pred

residuals2 <- sleepdatafinal$Quality.of.Sleep - predictions2$.pred

lasso_rmse <- sqrt(mean(residuals2^2))

sst2 <- sum((sleepdatafinal$Quality.of.Sleep - mean(sleepdatafinal$Quality.of.Sleep))^2)

ssr2 <- sum(residuals2^2)

lasso_r2 <- 1 - (ssr2 / sst2)

print(lasso_rmse)

print(lasso_r2)
```
The Observed vs. Predicted plot looks pretty good! Our RMSE value was once again quite low, at 0.0557 and our R-squared value was high at 0.997. Both of these metrics, combined with the Observed vs. Predicted plot indicate that the LASSO model is performing well on new data.

### 2024-05-01 products/manuscript/Manuscript.qmd
References Check

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "4-29-2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../manuscript/references_sleepdata.bib
editor: 
  markdown: 
    wrap: 72
---

>>>>

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "4-29-2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl 
editor: 
  markdown: 
    wrap: 72
---
